### 2012-06-23 trunk/ijcis-pos-model/sources/intro.tex
more modifications

A very specific type of information systems are database systems. A database contains data representing real objects or concepts. In real world, some aspects or properties of objects are time-variant or time-related. e.g., the moment of a bank transaction and the status of an employee in a company, are time-related and time-variant notions, respectively.
A temporal database schema is a database schema that models objects with time-related or time-variant properties. However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints and suitable ways of interactive with the human user.

For example, let us consider a hospital database containing data about the state of the patients in a given hospital. Two dates are stored: the date when the patient arrives to and leaves from the hospital. It is clear that a patient cannot leave the hospital if he or she has not arrived. Without further cautions, a hospital employee could insert a patient who is already in the hospital. A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.

A lot of research concerns temporal database models and their approaches to the modelling of time. The first efforts were towards the representation of historical information related to objects represented by records in a database~\cite{Clifford1985}. Some proposals tried to extend the Entity Relationship Model~\cite{Klopprogge1983}, without impact on any database standards like SQL~\cite{Sarda1990}.

@@

Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecision, vagueness, uncertainty or inconsistencies possibly contained in the descriptions of these temporal notions. These possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling. Consider as an example the description of the temporal notion in a sentence like `The Belfry of Bruges was finished between 01/01/1201 A.D. and 31/12/1300 A.D.'. These sentence contains imperfection because of the uncertainty in the used time-related expression. It is known that the building was finished on a single day, but this day is not precisely known.

>>>>

A very specific type of information systems are database systems. A database contains data representing real objects or concepts. In real world, some aspects or properties of objects are time-variant or time-related. e.g., the moment of a bank transaction and the status of an employee in a company, are time-related and time-variant notions, respectively.
A temporal database schema is a database schema that models objects with time-related or time-variant properties. However, the modelling of temporal aspects has a direct impact on the consistency of the temporal database, because the temporal nature of these aspects imposes extra integrity constraints and suitable ways of interaction with the human user.

% For example, let us consider an employee's database containing data about the positions of employees in a company. Two dates are stored: the starting and ending dates respectively of the contract. It is clear that an employee cannot have a contract with an ending date without an starting date. Without further cautions, a manager in the company could insert a new contract for an employee who is already working for the company. A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.

For example, let us consider a library database, modelling the presence of books in the library. Two dates are stored: the loan and the return date. It is clear that a book cannot be loaned again until it is returned. Without further cautions, a library employee could loan the same book several times even if it is not returned. A temporal database model will typically constrain record insertion and prevent similar modelling inconsistencies.

% let us consider a hospital database containing data about the state of the patients in a given hospital. Two dates are stored: the date when the patient arrives to and leaves from the hospital. It is clear that a patient cannot leave the hospital if he or she has not arrived. Without further cautions, a hospital employee could insert a patient who is already in the hospital. 
A lot of research concerns temporal database models and their approaches to the modelling of time. The first efforts were towards the representation of historical information related to objects represented by records in a database~\cite{Clifford1985}. Some proposals tried to extend the Entity Relationship Model~\cite{Klopprogge1983}, without impact on any database standards like SQL~\cite{Sarda1990}.

@@

Humans handle temporal information using certain temporal notions like time intervals or time points~\cite{Dyreson1994}, and they often have to deal with imperfections like imprecision, vagueness, uncertainty or inconsistencies possibly contained in the descriptions of these temporal notions. These possible imperfections in descriptions of temporal notions determine an important issue in temporal modelling. Consider as an example the description of the temporal notion in a sentence like `The Belfry of Bruges was finished between 01/01/1201 A.D. and 31/12/1300 A.D.'. This sentence contains imperfection because of the uncertainty in the used time-related expression. It is known that the building was finished on a single day, but this day is not precisely known.

### 2012-06-23 trunk/ijcis-pos-model/sources/preliminaries.tex
more modifications

\begin{example}
As an example, consider an employee's database. When an employee changes the status in the company, a record tracking the employee is stored. The exact time for this is not precisely known. Consider now that $X = \left[12, 2, 2\right]$ is the day of the month when an employee started a contract as a junior developer and $Y = \left[18, 2, 1 \right]$ is the day of the month when the employee changed the status to senior developer. It is possible to obtain $Z$ (the period of time between the junior and the senior contracts) by using a conjunctive combination of constraints.

\begin{align}
\nonumber
CC = \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace \\
\nonumber
Z = CC
\end{align}
$Z$ is a fuzzy interval defined by a trapezoidal shape given by $\left[12,\ 14,\ 18,\ 19 \right]$.
Figure \ref{fig:example-ill-known-by-const} illustrates the relations among the variables $X$, $Y$ and $Z$.
\end{example}

>>>>

\begin{example}
As an example, consider a historical database containing data about diplomatic medieval documents. The starting and ending dates when a diplomatic document was valid, are not precisely known.  Consider now that the time granularity are years. Then $X = \left[1112, 2, 2\right]$ is the starting year for the validity of a document. A new diplomatic document was valid in the year $Y = \left[1118, 2, 1 \right]$. Then, It is possible to obtain $Z$ (the period of time between the starting of both documents) by using a conjunctive combination of constraints.

\begin{align}
\nonumber
CC &= \left \lbrace C_1\left(>,X\right) \wedge C_2(\leq,Y) \right \rbrace \\
\nonumber
Z &= CC
\end{align}
$Z$ is a fuzzy interval defined by a trapezoidal shape given by $\left[1112,\ 1114,\ 1118,\ 1119 \right]$.
Figure \ref{fig:example-ill-known-by-const} illustrates the relations among the variables $X$, $Y$ and $Z$.
\end{example}
